Modular Model-Checking for Fault-Tolerant Distributed System Implementations

### 2016-12-02 doc/om-paper/paper.tex
Add \OMH macro

\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
\newcommand{\Id}{\ensuremath{\mathrm{Id}}}
\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}}

@@

To demonstrate the modeling technique, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}. The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm. In Section~\ref{sec:experimental}, we give performance metrics for the model. First, we benchmark the model, showing that despite the additional complexity, it scales farther than other model-checking verifications at the algorithmic level. The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{sal}. Developing invariants requires some user guidance. A user should not have to discover all new invariants for new protocols or implementations. We show that most of the lemmas required are not protocol specific and are modular, concerning only specific aspects of the system. To demonstrate this, we modify the OHM implementation in various ways to show that only local modifications are required to the model and proofs.

@@

To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid (OMH) algorithm~\cite{}, a more sophisticated version of the class Oral Messages (OM) algorithm. OM was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults. OMH was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults. However, OMH had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.

@@

\subsection{OMH(1) Algorithm}
OMH is a recursive algorithm that proceeds in rounds of communication. Let $OMH(m)$ be a recursive function implementing the algorithm, parameterized by the number of rounds, $m$. Consider a finite set of nodes $N$. Distinguish one node as the \emph{general}, $g$, and the remaining nodes $L = N \setminus \{g\}$ as the \emph{lieutenants}. We assume the identity of any general or lieutenant cannot be spoofed. Broadcast communication proceeds in rounds. Denote any message that is detectably faulty (e.g., fails a CRC) or is absent, by $ERR$. Additionally, in the algorithm, nodes report on values they have previously received. In doing so, nodes must differentiate \emph{reporting} $ERR$ from an $ERR$ itself. Let $R$ denote that an error is being reported. Finally, let $V$ be a special, designated value.

@@

\begin{itemize}
\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $ERR$).
\item {\bf $OM(m)$, $m > 0$}:
  \begin{enumerate}
  \item $g$ broadcasts a value to each lieutenant, $l$.
  \item\label{om:two} Let $l_v$ be the value received by $l \in L$ from $g$. Then for each $l$, execute $OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants. $l$ sends $l_v$ or $R$, if $l_v = ERR$.
  \item For each lieutenant $l \in L$, remove all $ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$. Compute the majority value over the remaining values, or $V$ if there is no majority. If the majority value is $R$, return $E$.
  \end{enumerate}
\end{itemize}

\noindent
In particular, $OMH(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.

OMH is designed to preserve \emph{validity} and \emph{agreement}
properties. Validity states that if the general is nonfaulty, then every lieutenant outputs the value sent by the general. More formally, let the general broadcast $v \neq ERR$, and let $l_i$ denote the output for lieutenant $i$:
%
\begin{equation}
  \tag{Validity}
    \forall \,i. \quad l_i = v
\end{equation}
%
Agreement states that each lieutenant outputs the same value. Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively:
%
\begin{equation}
  \tag{Agreement}
    \forall \,i, j. \quad l_i = l_j
\end{equation}
%

@@

\begin{figure}
  \centering
  \begin{tabular}{r||c|c|c|c|c|c|c|c|c|c|}
      \multicolumn{11}{c}{Receivers} \\
      \cline{2-11}
   Relays &   1  &  2  &   3  &  4  &   5  &  6   &  7   &  8  &  9  &  10 \\
      \hline \hline
      1   & 7    & 9   & 12   & 15  & 21   & 25   & 32   & 40  & 54  & 74  \\
      \hline
      2   & 17   & 14  & 21   & 30  & 42   & 53   & 74   & 99  & 144 & -   \\
      \hline
      3   & 21   & 22  & 40   & 50  & 81   & 102  & 155  & 279 & -   & -   \\
      \hline
      4   & 27   & 34  & 59   & 99  & 141  & 237  & 1114 & -   & -   & -   \\
      \hline
      5   & 22   & 94  & 125  & 335 & TO   & 1406 & -    & -   & -   & -   \\
      \hline
      6   & 36   & 132 & 2966 & 844 & 2457 & -    & -    & -   & -   & -   \\
      \hline
      7   & 83   & 487 & TO   & TO  & -    & -    & -    & -   & -   & -   \\
      \hline
      8   & 298  & TO  & TO   & -   & -    & -    & -    & -   & -   & -   \\
      \hline
      9   & 1428 & TO  & -    & -   & -    & -    & -    & -   & -   & -   \\
      \hline
     10   & TO   & -   & -    & -   & -    & -    & -    & -   & -   & -   \\
      \hline
  \end{tabular}
  \label{fig:benchmark}
  \caption{Benchmark of full proof computation time for OMH(1) implementation. Times are in seconds with a timeout (TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
\end{figure}

@@

Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults~\cite{om1-synth}. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. That said, their model of OM(1) (not OMH(1)), timeouts after 10~hours, and we assume a voting algorithm is not explicitly specified. \lee{for what size? Also, Rushby's model?} Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.

>>>>

\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
\newcommand{\OMH}{\ensuremath{\mathrm{OMH}}}
\newcommand{\Id}{\ensuremath{\mathrm{Id}}}
\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}}

@@

To demonstrate the modeling technique, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}. The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm. In Section~\ref{sec:experimental}, we give performance metrics for the model. First, we benchmark the model, showing that despite the additional complexity, it scales farther than other model-checking verifications at the algorithmic level. The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{sal}. Developing invariants requires some user guidance. A user should not have to discover all new invariants for new protocols or implementations. We show that most of the lemmas required are not protocol specific and are modular, concerning only specific aspects of the system. To demonstrate this, we modify the OHM implementation in various ways to show that only local modifications are required to the model and proofs.

@@

To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid ($\OMH$) algorithm~\cite{csl-93-2}, a more sophisticated version of the classic Oral Messages ($\OM$) algorithm. $\OM$ was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults. $\OMH$ was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults. However, $\OMH$ had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.

@@

\ben{The word "model" is very overloaded in the paper so far. I wonder if we could use a different word like "modeling framework", or something more concise, for the one conotation and just "model" for the concrete mathematical description of a particular system?}

\subsection{$\OMH(m)$ Algorithm}
$\OMH$ is a recursive algorithm that proceeds in rounds of communication. Let $\OMH(m)$ be a recursive function implementing the algorithm, parameterized by the number of rounds, $m$. Consider a finite set of nodes $N$. Distinguish one node as the \emph{general}, $g$, and the remaining nodes $L = N \setminus \{g\}$ as the \emph{lieutenants}. We assume the identity of any general or lieutenant cannot be spoofed. Broadcast communication proceeds in rounds. Denote any message that is detectably faulty (e.g., fails a CRC) or is absent, by $ERR$. Additionally, in the algorithm, nodes report on values they have previously received. In doing so, nodes must differentiate \emph{reporting} $ERR$ from an $ERR$ itself. Let $R$ denote that an error is being reported. Finally, let $V$ be a special, designated value.

@@

\begin{itemize}
\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $ERR$).
\item {\bf $OM(m)$, $m > 0$}:
  \begin{enumerate}
  \item $g$ broadcasts a value to each lieutenant, $l$.
  \item\label{om:two} Let $l_v$ be the value received by $l \in L$ from $g$. Then for each $l$, execute $\OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants. $l$ sends $l_v$ or $R$, if $l_v = ERR$.
  \item For each lieutenant $l \in L$, remove all $ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$. Compute the majority value over the remaining values, or $V$ if there is no majority. If the majority value is $R$, return $E$.
  \end{enumerate}
\end{itemize}

\noindent
In particular, $\OMH(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.

$\OMH$ is designed to preserve \emph{validity} and \emph{agreement}
properties. Validity states that if the general is nonfaulty, then every lieutenant outputs the value sent by the general. More formally, let the general broadcast $v \neq ERR$, and let $l_i$ denote the output for lieutenant $i$:
%
\begin{equation}
  \tag{Validity}
    \forall \,i. \quad l_i = v
\end{equation}
%
Agreement states that each lieutenant outputs the same value. Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively:
%
\begin{equation}
  \tag{Agreement}
    \forall \,i, j. \quad l_i = l_j
\end{equation}
%

@@

\begin{figure}
  \centering
  \begin{tabular}{r||c|c|c|c|c|c|c|c|c|c|}
      \multicolumn{11}{c}{Receivers} \\
      \cline{2-11}
   Relays &   1  &  2  &   3  &  4  &   5  &  6   &  7   &  8  &  9  &  10 \\
      \hline \hline
      1   & 7    & 9   & 12   & 15  & 21   & 25   & 32   & 40  & 54  & 74  \\
      \hline
      2   & 17   & 14  & 21   & 30  & 42   & 53   & 74   & 99  & 144 & -   \\
      \hline
      3   & 21   & 22  & 40   & 50  & 81   & 102  & 155  & 279 & -   & -   \\
      \hline
      4   & 27   & 34  & 59   & 99  & 141  & 237  & 1114 & -   & -   & -   \\
      \hline
      5   & 22   & 94  & 125  & 335 & TO   & 1406 & -    & -   & -   & -   \\
      \hline
      6   & 36   & 132 & 2966 & 844 & 2457 & -    & -    & -   & -   & -   \\
      \hline
      7   & 83   & 487 & TO   & TO  & -    & -    & -    & -   & -   & -   \\
      \hline
      8   & 298  & TO  & TO   & -   & -    & -    & -    & -   & -   & -   \\
      \hline
      9   & 1428 & TO  & -    & -   & -    & -    & -    & -   & -   & -   \\
      \hline
     10   & TO   & -   & -    & -   & -    & -    & -    & -   & -   & -   \\
      \hline
  \end{tabular}
  \label{fig:benchmark}
  \caption{Benchmark of full proof computation time for $\OMH(1)$ implementation. Times are in seconds with a timeout (TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
\end{figure}

@@

Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults~\cite{om1-synth}. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. That said, their model of $\OM(1)$ (not $\OMH(1)$), timeouts after 10~hours, and we assume a voting algorithm is not explicitly specified. \lee{for what size? Also, Rushby's model?} Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
